Estimating Neighborhood Rents with Scraped Data

### 2018-07-27 manuscript/index.Rmd
adding section break tags after models, results
Small typo-type edits.

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "../footer.html"
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

@@

Figure 2 is a heatmap for the time-series of one-bedroom listings per day. The grey cells indicate days where our scraper was offline. Outages were more common in the earliest phases of work, but now are relatively rare.

@@

In the following graphic, the left most pane shows median one-bedroom asking rent estimates using data collected from Craigslist in 2017-2018. The middle pane shows Dupre+Scott tract median 1B contract rent estimates for one-bedrooms in 20+ unit apartment structures according to their September 2017 (Fall) survey of local landlords and property managers. These data were only available in 2000 Census geography, so the neighborhood units differ slightly (124 tracts versus 134 tracts) in this map compared to the other two. Though less than ideal, this limitation does not drastically alter the comparison between data.  Finally, on the right are ACS 2012-2016 5-Year Estimates (the most recent ACS small-area data product available). In all map panes, grey cells denote missing values.

@@

Figure 5 shows a bubbleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e. about 2/3 of the 124 tracts in Seattle using 2000 definitions). The purple dashed line indicates the loess smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

![Bubbleplot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties. Purple dashed line indicates loess fit.](../output/rent_scatter_across_1B.png)

@@

The space and time smoothing models used in this section partially pool the data (e.g. the same observation at different points in time, a tract's neighboring units) to measure the degree of variation across different quarters and at different points in space. These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability). Modelling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area). The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: "header.html"
      after_body: "../footer.html"
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investigates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

@@

Figure 2 is a heatmap for the time-series of one-bedroom listings per day. The gray cells indicate days where our scraper was offline. Outages were more common in the earliest phases of work, but now are relatively rare.

@@

In the following graphic, the left most pane shows median one-bedroom asking rent estimates using data collected from Craigslist in 2017-2018. The middle pane shows Dupre+Scott tract median 1B contract rent estimates for one-bedrooms in 20+ unit apartment structures according to their September 2017 (Fall) survey of local landlords and property managers. These data were only available in 2000 Census geography, so the neighborhood units differ slightly (124 tracts versus 134 tracts) in this map compared to the other two. Though less than ideal, this limitation does not drastically alter the comparison between data.  Finally, on the right are ACS 2012-2016 5-Year Estimates (the most recent ACS small-area data product available). In all map panes, gray cells denote missing values.

@@

Figure 5 shows a bubbleplot of 2000-definition Census tract one-bedroom rent estimates between Craigslist and Dupre+Scott for the 80 tracts where comparable estimates exist (i.e. about 2/3 of the 124 tracts in Seattle using 2000 definitions). The purple dashed line indicates the LOESS smoother fit between the measures. Looking at the measure comparison, the first takeaway is that there is strong correlation ($r = .82$) and relative agreement between the measures (mean absolute difference = $186). The Craigslist measure reflects asking rents and the Dupre+Scott measure reflects contract rents for a neighborhood (i.e. the asking rent less an incentives a landlord offers). Given the currently low vacancy rates in most properties, the only neighborhoods where there would theoretically be more salience to the asking vs contract rent distinction (i.e. a wider gap in these indicators) are at the high end of the distribution where demand for luxury apartments has slackened some. At the high-end of the neighborhood rent distribution, however, Craigslist and Dupre+Scott are in arguably the strongest agreement.

@@

![Bubbleplot of neighborhood 1-bedroom rents across Craigslist and Dupre+Scott. Craigslist represents median asking rent among all one-bedroom listings, Dupre+Scott consists of median contract rents among one-bedroom units in 20+ unit properties. Purple dashed line indicates LOESS fit.](../output/rent_scatter_across_1B.png)

@@

The space and time smoothing models used in this section partially pool the data (e.g. the same observation at different points in time, a tract's neighboring units) to measure the degree of variation across different quarters and at different points in space. These models are common in epidemiology and disease-mapping applications where researchers need to estimate a quantity of interest like a relative risk but have sparse data for some areas (which may differ from truth due to sampling variability). Modeling a panel's error structure with random-effects for space and time penalizes bumpiness (local deviations) in predicted values and theoretically improves out-of-sample accuracy (because large local deviations can stem from low $n$ in a given area). The spatial models in particular allow for missing data to be imputed based off of a local average of neighbors rather than the global or period-specific average.

### 2018-07-27 manuscript/manuscript.Rmd
adding section break tags after models, results
Small typo-type edits.

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
output:  pdf_document
abstract: "Online classified ads are a key source of information about local rental housing markets. This research first assesses the agreement of census tract rent estimates based on scraped rental advertisements with existing data sources. The second component of the paper investivates temporal and spatial dynamics in one-bedroom rents among Seattle neighborhoods using spatial models. Posterior distributions of all model hyperparameters and parameters are estimated with integrated nested Laplace approximations (INLA). Training data for models cover quarterly tract estimates for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed. The top model assumes that spatial and temporal dynamics are separable, i.e. independent of each other, and non-interactive. This spatial random-effects model with AR(1) temporal structure achieves a mean absolute error (MAE) of about $100 for the training data and $140 for the test data."
---

>>>>

---
title: "Estimating Neighborhood Rents with Scraped Data"
author: "Chris Hess and Sarah Chasins"
output:
  html_document:
    df_print: paged
abstract: Online classified ads are a key source of information about local rental
  housing markets. This research first assesses the agreement of census tract rent
  estimates based on scraped rental advertisements with existing data sources. The
  second component of the paper investivates temporal and spatial dynamics in one-bedroom
  rents among Seattle neighborhoods using spatial models. Posterior distributions
  of all model hyperparameters and parameters are estimated with integrated nested
  Laplace approximations (INLA). Training data for models cover quarterly tract estimates
  for Q2 2017 through Q1 2018, while the test data are Q2 2018 median values as observed.
  The top model assumes that spatial and temporal dynamics are separable, i.e. independent
  of each other, and non-interactive. This spatial random-effects model with AR(1)
  temporal structure achieves a mean absolute error (MAE) of about $100 for the training
  data and $140 for the test data.
---
